Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-01-18 numerical-mhd-ct.tex
Adding MHD-CT text, references

\subsection{MHD: Constrained transport}
\label{sec.num.mhd-ct}

>>>>

\subsection{MHD: Constrained transport}
\label{sec.num.mhd-ct}
\def\Bvec{{\bf B}}
\def\Bf{Bf}
\def\Bc{Bc}
\def\Evec{{\bf E}}
\def\Divb{\ensuremath{\nabla \cdot \Bvec}}
Physically, the divergence of the magnetic field, \Divb, is
identically zero due to the fact that the evolution of the magnetic field is the
curl of a vector, and the divergence of the curl of a vector is identically
zero:
\begin{eqnarray}
\partial \Bvec/\partial t & =& \frac{1}{a} \nabla \times \Evec\\
\Divb & =& 0.
\end{eqnarray}
The Constrained Transport (CT) method \citep{Evans88, Balsara99} of MHD evolution employs this same vector
property to evolve the magnetic field in a manner that preserves $\Divb=0$ to
the ability of the computer.  The electric field is computed from the fluxes
from the Riemann solver.  The curl of that electric field is then used to update
the  magnetic field.   The advantage of this method is that it preserves \Divb\
to machine precision.  The drawback is twofold: increased algorithm  complexity,
as will be described in this section; and since only the update of the magnetic
field is divergence free, any monopoles created by other numerical sources
persist.

Enzo MHD-CT has been described in horrid detail in \citet{Collins10}, so the
description here will be an overview only.  The base Godunov method, described
in \citet{Li08a}, uses spatially and temporally second order reconstruction
(both MUSCL-Hancock and Piecewise Linear Method), a selection of Riemann solvers
including HLLC and HLLD \citep{Mignone07}.  The CT methods are the first order
method described by \citet{Balsara99} and the second order methods described in
\citet{Gardiner05}.  The AMR machinery is described by \citet{Balsara01} and
\citet{Collins10}.

The increased complexity comes in the form of area-averaged face and
length-averaged edge centered variables, while the rest of Enzo employs
predominantly volume-averaged cell centered variables.  The magnetic field is
represented by both a face centered field, \Bf, and a cell centered field, \Bc.
The electric field is edge centered.
The magnetic field is updated in four steps:  first, the
Riemann problem is solved in the traditional manner, using the cell centered
field; second, an edge-centered
electric field is computed using the fluxes from the Riemann solver;  third, the curl
of that electric field is used to update the face centered field; finally,
the cell centered magnetic field is updated with an average of the face centered
field.

Divergence free AMR is slightly more complex than the AMR employed elsewhere in
Enzo.  First, the interpolation must be constrained to be divergence free.
Thus, all three face-centered field components are interpolated in concert.
Second, any magnetic field information from previous time steps must be included
in the interpolation, making the interpolation more complex than the simple
parent-child relation used for other fields.  Third, the flux correction
involves more possible grid relations than traditional AMR.  In order to
circumvent this last complexity, the electric field is projected from fine grids
to parent grids, rather than the magnetic field, which is then used to re-update
the parent magnetic field.  This is described in detail in \citet{Balsara99,
Collins10}

Enzo's MHD-CT has been used in a variety of contexts, from galaxy cluster
evolution \citep{Xu11,Xu12}, synthetic observations of radio halos
\citep{Skillman13},  primordial star formation \citep{Xu08}, 
present day star formation \citep{Collins11,Collins12}.
